---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(bayesTPC)
```

# bayesTPC: An accessible interface to fit Bayesian thermal performance curves in R.

<!-- badges: start -->
[![R-CMD-check](https://github.com/johnwilliamsmithjr/bayesTPC/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/johnwilliamsmithjr/bayesTPC/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/johnwilliamsmithjr/bayesTPC/branch/master/graph/badge.svg)](https://app.codecov.io/gh/johnwilliamsmithjr/bayesTPC?branch=master)
<!-- badges: end -->

**bayesTPC** is an R package to help conveniently fit common thermal performance models using Bayesian MCMC. We provide functionality to substitute for or supplement BUGS modelling workflows, especially for those without experience in Bayesian methods. To this end, **bayesTPC** contains a simple model specification format, BUGS configuration helpers, end-to-end model fitting methods, and model diagnostics.

For a detailed introduction, visit [this](https://seansorek.github.io/bayesTPC/pages/intro.html) tutorial.

## Installation

You can install the development version of **bayesTPC** from [GitHub](https://github.com/) with:

``` r
remotes::install_github("johnwilliamsmithjr/bayesTPC")
```

## Functionality

### Models

`bayesTPC` offers many pre-configured TPC models. We can use `get_models()` to view all implemented models.
```{r}
get_models()
```

To see the exact specification of a particular model, we can use
```{r}
get_default_model_specification("briere")
```

Use `get_formula()`, `get_default_priors()`, and `get_default_constants()` to display the individual components of a any model included in `bayesTPC`.

For those with experience in fitting models in BUGS, `configure_model()` provides pre-formatted code.

### Fitting

The fitting workhorse of `bayesTPC` is the `b_TPC()` function. With `b_TPC()`, a user only needs to specify the dataset and desired model and `bayesTPC` handles configuring and running an MCMC using `nimble`. The user can also change the model priors, the sampler used, and other hyperparameters of the MCMC.

### Changing and Specifying models

One can also create their own models by specifying a formula and priors using `specify_model()`, along with its wrappers `specify_normal_model()` and `specify_binomial_model()`.

`bayesTPC` accepts both the name of a model and the model object itself as inputs to its workhorse functions. This provides two methods for users to customize existing models: providing customized priors and constants in the options of the functions themselves, or by storing the model object in the local environment and using the `change_priors()` and `change_constants()` functions. Note that since R objects are immutable, the modified models must be stored again.

### Summaries

The fit object returned by `b_TPC()` can be summarized in multiple ways. `summary()` gives a detailed summary of the MCMC results, and `plot()` shows the fit given by the center (mean or median) and bounding (95% quantiles or hdi) MCMC samples. Similarly, `posterior_predictive()` and `plot_prediction()` take samples of the posterior to predict new values.

### Diagnostics

`bayesTPC` provides multiple MCMC diagnostic plots. `traceplot()` wraps `coda::traceplot()` and shows the sampled values by sequential iteration. `bayesTPC_ipairs()` wraps `IDPmisc::ipairs()` and shows the pairwise joint posterior distributions of all model parameters. These wrappers are provided for convenience.

The `ppo_plot()` function shows the overlap between the priors specified versus a kernel density estimation of the posterior sample.

### Vectraits API

`bayesTPC` also includes a basic API to access the Vectraits database of disease vector trait data. One can either retrieve a pre-known dataset using `get_VB_dataset()` or `get_VB_datasets()`, or search specific keywords using `find_VB_datasets()`.

